---
title: "UFA&Visualizations"
author: "Ingrid Backman"
date: "2024-09-26"
output: html_document
---
#UFA and some related collocate visualizations

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#code from lancaster, somehow mine might differ a bit?

#Calculate Gwet's AC1; b...input data frame
#i = 1; v <- c(); while(i+1 < ncol(b)) {n=(gwet.ac1.raw(b[,i:(i+2)])[3]);v<- c(v, n); i= i+1; }

#Prepare data frame
#h<-seq(from, to, by = 1); g<-data.frame(h,v)

#Produce graph
#p<-ggplot(g, aes(x = g[,1], y =g[,2])) + xlim(from, to)+ scale_x_continuous(breaks = seq(from, to, by = 10)) + geom_point()  + xlab("Time") + ylab("AC1"); p + stat_smooth(method = "gam", formula = y ~ s(x, bs = "cr", fx=FALSE, k =10), size = 1, fill="#707070", level = 0.95 )+ stat_smooth(method = "gam", formula = y ~ s(x, bs = "cr", fx=FALSE, k =10), size = 1, fill="#FFFF00", level = 0.99)
```


#clunky claude refactored version of what i was doing earlier more manually
```{r old ver}
library(fs)
library(tidyverse)
library(irrCAC)
library(pacman)

# Main function to process a single folder of collocate data
process_collocate_folder <- function(folder_path, 
                                   mi_score_threshold = 3,
                                   output_dir = ".") {

  folder_info <- extract_folder_info(folder_path)
  
  dir.create(output_dir, showWarnings = FALSE, recursive = TRUE)
  
  target_word <- ifelse(grepl("FACTS", folder_path), "facts", 
                      ifelse(grepl("FACT", folder_path), "fact", "unknown"))
  
  window = folder_info$window
  
  all_data <- load_collocate_data(folder_path, mi_score_threshold)
  spread_data <- prepare_spread_data(all_data)
  
  output_file_base <- paste0(target_word, "_collocates_",
                            folder_info$start_year, "_",
                            folder_info$end_year,
                            folder_info$window, "")
  
  save_spread_data(spread_data, output_file_base, output_dir)
  
  # Calculate Gwet's AC1
  g <- calculate_gwet_ac1(spread_data)
  
  create_ufa_plot(g, output_file_base, output_dir, target_word, window)
  
  return(list(
    all_data = all_data,
    spread_data = spread_data,
    gwet_ac1 = g,
    folder_info = folder_info
  ))
}

# Helper function to extract information from folder name
extract_folder_info <- function(folder_path) {
  # Get the parent folder name instead of just the last directory
  folder_name <- basename(dirname(folder_path))
  
  year_range <- str_extract(folder_name, "\\d{4}-\\d{4}")
  
  if (!is.na(year_range)) {
    years <- strsplit(year_range, "-")[[1]]
    start_year <- years[1]
    end_year <- years[2]
  } else {
    warning("Could not find year range in folder name: ", folder_name)
    start_year <- "UNKNOWN"
    end_year <- "UNKNOWN"
  }
  
  # Fixed window pattern to correctly extract the number after 'w'
  window <- str_extract(folder_name, "w\\d")
  if (is.na(window)) {
    warning("Could not find window size in folder name: ", folder_name)
    window <- "3"  # Default to 3 if not found
  } else {
    window <- substr(window, 2, 2)  # Extract just the number
  }
  
  # Update word count extraction to handle both formats
  word_count <- ifelse(
    grepl("css", folder_name),
    "css",
    str_extract(folder_name, "\\d+[-_]\\d+")
  )
  
  # Add some debugging output
  message("Extracted from folder: ", folder_name)
  message("Window size: ", window)
  
  return(list(
    start_year = start_year,
    end_year = end_year,
    window = window,
    word_count = word_count
  ))
}

# Function to process individual collocate files
process_file <- function(file_path) {
  year <- as.numeric(str_extract(basename(file_path), "\\d{4}"))
  lines <- readLines(file_path)
  
  if (length(grep("^\\d+\\s+\\w+", lines)) == 0) {
    return(data.frame(word = character(), mi_score = numeric(), year = numeric()))
  }
  
  collocates_data <- lines[grep("^\\d+\\s+\\w+", lines)]
  
  collocates <- lapply(collocates_data, function(line) {
    parts <- strsplit(line, "\\s+")[[1]]
    list(
      word = parts[2],
      mi_score = as.numeric(parts[7])
    )
  })
  
  df <- do.call(rbind, lapply(collocates, as.data.frame))
  df$year <- year
  
  return(df)
}

# Load and filter collocate data
load_collocate_data <- function(data_dir, mi_score_threshold = 3) {
  collocate_files <- fs::dir_ls(data_dir, regexp = "\\.txt$")
  
  all_data <- do.call(rbind, lapply(collocate_files, process_file))
  all_data <- all_data %>%
    distinct(word, year, .keep_all = TRUE) %>%
    filter(mi_score >= mi_score_threshold) %>%
    mutate(present = 1)
  
  return(all_data)
}

# Remaining helper functions
prepare_spread_data <- function(all_data) {
  spread_data <- all_data %>%
    pivot_wider(names_from = year, values_from = present, values_fill = 0)
  
  return(spread_data)
}

save_spread_data <- function(spread_data, base_filename, output_dir = ".") {
  output_file <- file.path(output_dir, paste0(base_filename, ".csv"))
  write.table(spread_data, file = output_file, sep = "\t")
}

calculate_gwet_ac1 <- function(data) {
  print(data)
  i <- 3 # Skipping the first two columns (collocate and mi score)
  v <- c()
  years <- c()
  
  while (i + 1 < ncol(data)) { 
    if (sum(colSums(data[, i:(i+2)]) > 0) >= 1) {  
      n <- gwet.ac1.raw(data[, i:(i+2)])$est$coeff.val 
      v <- c(v, n)
      years <- c(years, as.numeric(colnames(data)[i]))
    } else {
      message("Skipping AC1 calculation for years ", colnames(data)[i], "-", colnames(data)[i+2], " due to insufficient data")
    }
    i <- i + 1
  }
  
  return(data.frame(year = years, ac1 = v))  # Return a data frame with the years and ac1 values
}


create_ufa_plot <- function(g, base_filename, output_dir = ".", target_word, window_type) {
  output_file <- file.path(output_dir, paste0(base_filename, "_ufa.png"))
  
  # Validate input data
  if (nrow(g) == 0 || !all(c("year", "ac1") %in% names(g))) {
    stop("Invalid or empty data provided to create_ufa_plot")
  }
  
  # Ensure numeric values
  g$year <- as.numeric(g$year)
  g$ac1 <- as.numeric(g$ac1)
  
  # Remove any NA values
  g <- g[complete.cases(g), ]
  
  if (nrow(g) == 0) {
    stop("No valid data points after cleaning")
  }
  
  # Calculate proper breaks and limits
  min_year <- floor(min(g$year, na.rm = TRUE))
  max_year <- ceiling(max(g$year, na.rm = TRUE))
  break_seq <- seq(min_year, max_year, by = 20)
  # Ensure the last year is included if it's not already in the sequence
  if (max_year > max(break_seq)) {
    break_seq <- c(break_seq, max_year)
  }
  
  p <- ggplot(g, aes(x = year, y = ac1)) + 
    geom_point() + 
    scale_x_continuous(
      breaks = break_seq,
      limits = c(min_year, max_year),
      expand = expansion(mult = 0.02)  # Add small padding
    ) +
    xlab("Year") + 
    ylab("AC1") + 
    ggtitle(
      bquote(
        "Results of UFA for " * italic(.(target_word)) *
          ": 3a-MI(3), " * .(window_type) * "L-" * .(window_type) * "R, " *
          "C3"[relative] * "-" *
          "NC3"[relative]
      )
    )
  
  final_plot <- p + 
    stat_smooth(method = "gam", 
                formula = y ~ s(x, bs = "cr", fx = FALSE, k = 10), 
                size = 1, fill = "#707070", level = 0.95) + 
    stat_smooth(method = "gam", 
                formula = y ~ s(x, bs = "cr", fx = FALSE, k = 10), 
                size = 1, fill = "#FFFF00", level = 0.995) +
    theme_minimal(base_size = 15) + 
    theme(
      panel.background = element_rect(fill = "white"),  
      plot.background = element_rect(fill = "white"),
      axis.text.x = element_text(angle = 45, hjust = 1),
      plot.title = element_text(margin = margin(b = 10)),
      axis.title.x = element_text(margin = margin(t = 10)),
      axis.title.y = element_text(margin = margin(r = 10))
    )
  
  ggsave(output_file, plot = final_plot, width = 12, height = 8, units = "in", dpi = 300)
}
```
#this is kind of clumsy naming and calling, can fix it later
#also some redundant variables etc
```{r}
#3 word windows
# results1 <- process_collocate_folder(
#   folder_path = "../collocation_results/FACT/collocate_results_1665-1958_FACT_css3_w3/collocation_txt",
#   output_dir = "collocate_results/3window"
# )
# 
# results2 <- process_collocate_folder(
#   folder_path = "../collocation_results/FACTS/collocate_results_1665-1958_FACTS_css3_w3/collocation_txt",
#   output_dir = "collocate_results/3window"
# )

#5 word windows
results3 <- process_collocate_folder(
  folder_path = "../collocation_results/FACT/collocate_results_1665-1958_FACT_css3_w5/collocation_txt",
  output_dir = "collocate_results/5window"
)

#this is old, css=3
results4 <- process_collocate_folder(
  folder_path = "../collocation_results/FACTS/collocate_results_1665-1958_FACTS_css3_w5/collocation_txt",
  output_dir = "collocate_results/5window"
)
```


```{r}
#5 word window latest (2 scaling for facts); too many? its certainly a choice not to use this one but they dont differ in the UFA plot all that much anyway
 # results4 <- process_collocate_folder(
 #   folder_path = "../collocation_results/FACTS/collocate_results_1665-1958_FACTS_css2_w5/collocation_txt",
 #   output_dir = "collocate_results/5window"
 # )
```

#plotting the plots for easier visual comparison

```{r}
#3l3r
library(knitr)

folder_to_ignore <- c("5window", "old/5window", "old/3window")

files <- list.files("collocate_results/", 
                    pattern = "_ufa\\.png$", 
                    full.names = TRUE, 
                    recursive = TRUE)

files <- files[!grepl(paste(folder_to_ignore, collapse = "|"), files) & grepl("3window", files)]

if (length(files) == 0) {
  print("No images found with '_ufa.png' in the filename.")
} else {
  include_graphics(files)
}


#5l5r
folder_to_ignore <- c("3window", "old/5window", "old/3window")

files <- list.files("collocate_results/", 
                    pattern = "_ufa\\.png$", 
                    full.names = TRUE, 
                    recursive = TRUE)

files <- files[!grepl(paste(folder_to_ignore, collapse = "|"), files) & grepl("5window", files)]

# Check if there are files left, then plot or print a message
if (length(files) == 0) {
  print("No images found with '_ufa.png' in the filename.")
} else {
  include_graphics(files)
}



```

```{r}
library(dplyr)
library(tidyr)

create_collocate_table <- function(results, target_word = "fact") {
  collocate_table <- results$all_data %>%
    filter(mi_score >= 3) %>%  # Filter for MI scores >= 3
    arrange(year) %>%  # Sort by MI score first, then year
    select(Year = year,
           Collocate = word,
           `MI Score` = mi_score) %>%
    as_tibble()
  
  return(collocate_table)
}

#fact
#r1 = create_collocate_table(results1) #fact 3L3R
r3= create_collocate_table(results3) #fact 5L5R

#facts
#r2 = create_collocate_table(results2) #facts 3L3R
r4 = create_collocate_table(results4) #facts 5L5R


#r1
r3
#r2
r4

#write.csv(r1, "collocate_results/dfs/fact3.csv")
write.csv(r3, "collocate_results/dfs/fact5.csv")

#write.csv(r2, "collocate_results/dfs/facts3.csv")
write.csv(r4, "collocate_results/dfs/facts5.csv")
```


#PLOTTING, COULD BE SOMEWHERE ELSE
```{r}
#from claude, just for funsies (illustrates how few words are used for facts)
library(dplyr)
library(tidyr)
library(ggplot2)

# Function to count unique collocates per year for each dataset
analyze_unique_collocates <- function(results_list, labels) {
  # Create a list to store counts for each dataset
  yearly_counts <- list()
  
  for(i in seq_along(results_list)) {
    counts <- results_list[[i]] %>%
      group_by(Year) %>%
      summarize(unique_words = n_distinct(Collocate)) %>%
      mutate(dataset = labels[i])
    
    yearly_counts[[i]] <- counts
  }
  
  # Combine all counts
  all_counts <- bind_rows(yearly_counts)
  
  # Create visualization
  ggplot(all_counts, aes(x = Year, y = unique_words, color = dataset)) +
    geom_line(size = 1) +
    geom_point() +
    theme_minimal() +
    labs(title = "Unique Collocates Over Time",
         subtitle = "Comparison between 'fact' and 'facts', PMI >=3",
         y = "Number of Unique Collocates",
         color = "Dataset") +
    theme(legend.position = "bottom",
          plot.title = element_text(face = "bold"),
          axis.title = element_text(face = "bold"))
}

# Create list of your results
results_list <- list(
  create_collocate_table(results3),
  create_collocate_table(results4)
)

# Labels for each dataset
labels <- c( "fact 5L5R", "facts 5L5R")

# Generate the visualization
analyze_unique_collocates(results_list, labels)

# Print summary statistics
for(i in seq_along(results_list)) {
  cat("\nSummary for", labels[i], ":\n")
  print(summary(results_list[[i]] %>% 
                group_by(Year) %>% 
                summarize(unique_words = n_distinct(Collocate)) %>%
                pull(unique_words)))
}
```

#this plot is kept just for comparison, the version used in the paper is in extra_freq_vis
```{r}
library(ggplot2)
library(dplyr)

# Function for detailed visualization with MI scores
create_detailed_collocate_plot <- function(data, 
                                         target_word = "fact/facts", 
                                         window_size = "3L3R",
                                         save_plot = TRUE,
                                         output_dir = "collocate_results") {
  # Define subdirectory based on window size
  subdirectory <- ifelse(window_size == "3L3R", "3window", "5window")
  output_path <- file.path(output_dir, subdirectory)
  
  # Ensure output directory exists
  if (!dir.exists(output_path)) {
    dir.create(output_path, recursive = TRUE)
  }
  
  filtered_data <- data %>%
    filter(mi_score >= 3)
  
  # Create the plot
  detailed_plot <- filtered_data %>%
    arrange(year, desc(mi_score)) %>%
    ggplot(aes(x = year, y = reorder(word, mi_score))) +
    geom_point(aes(size = mi_score, color = mi_score)) +
    scale_color_viridis_c(option = "plasma") +
    scale_size_continuous(range = c(2, 6)) +
    theme_bw(base_size = 12) +  # Changed to theme_bw for white background
    labs(
      title = paste("Distribution of Collocates for '", target_word, "' Over Time (", window_size, ")", sep=""),
      x = "Year",
      y = "Collocate",
      size = "MI Score",
      color = "MI Score"
    ) +
    theme(
      plot.title = element_text(size = 14, face = "bold"),
      axis.text.y = element_text(size = 8),
      legend.position = "right",
      panel.grid.major = element_line(color = "gray90"),
      panel.grid.minor = element_line(color = "gray95"),
      panel.background = element_rect(fill = "white", color = NA)  # Ensures background is white
    ) +
    scale_x_continuous(
      breaks = seq(min(filtered_data$year), 
                  max(filtered_data$year), 
                  by = 20)
    )
  
  # Save the plot if requested
  if(save_plot) {
    filename <- file.path(output_path, 
                         paste0(target_word, "_collocate_distribution_", 
                               window_size, ".png"))
    ggsave(filename, detailed_plot, width = 15, height = 10, dpi = 300)
  }
  
  return(detailed_plot)
}

# Example usage for all your datasets:
# For fact 3L3R
#p1_detailed <- create_detailed_collocate_plot(results1$all_data, "fact", "3L3R", save_plot = F) #the one from frequency is used

# For facts 3L3R
#p2_detailed <- create_detailed_collocate_plot(results2$all_data, "facts", "3L3R", save_plot = F)

# For fact 5L5R
p3_detailed <- create_detailed_collocate_plot(results3$all_data, "fact", "5L5R", save_plot = F)

# For facts 5L5R
p4_detailed <- create_detailed_collocate_plot(results4$all_data, "facts", "5L5R", save_plot = F)

```


```{r}
#print(p1_detailed)
#print(p2_detailed)
```
```{r}
print(p3_detailed)
print(p4_detailed)
```
```{r}
#find all unique words in word
results3$all_data %>%
  filter(mi_score >= 3) %>%
  group_by(word) %>%
  summarize(max_mi_score = max(mi_score)) %>%  # Get the highest MI score for each word
  arrange(desc(max_mi_score)) %>%  # Sort by MI score in descending order
  ungroup()
```

```{r}
results4$all_data %>%
  filter(mi_score >= 3) %>%
  group_by(word) %>%
  summarize(max_mi_score = max(mi_score)) %>%  # Get the highest MI score for each word
  arrange(desc(max_mi_score)) %>%  # Sort by MI score in descending order
  ungroup()
```
#minimum ac1 for potential reporting needs
```{r}
results3$gwet_ac1 %>%
  filter(ac1 == min(ac1))

results4$gwet_ac1 %>%
  filter(ac1 == min(ac1))
```

